---
title: "Machine Learning Final Project"
author: "Lazar Kirven"
date: "December 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. This project will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants and predict the manner in which they did the exercise.


## Data Exploratory Section

```{r}
library(caret); library(randomForest);
mytmpdir = tempdir()
tempTrain <- tempfile(tmpdir = mytmpdir)
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", tempTrain)
rawTrainData <- read.csv(tempTrain, header = TRUE, sep = ",", na.strings = c("NA", ""))

tempTest <- tempfile(tmpdir = mytmpdir)
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", tempTest)
rawTestData <- read.csv(tempTest, header = TRUE, sep = ",", na.strings = c("NA", ""))
unlink(mytmpdir)
```


## Data Cleaning

Below is some code that removes any column that just contains NAs:

```{r]}
badColumns <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")
cleanTrainData <- rawTrainData[, colSums(is.na(rawTrainData)) == 0]
# bad columns
cleanTrainData <- cleanTrainData[, -c(1:7)]
str(cleanTrainData)
cleanTestData <- rawTestData[, colSums(is.na(rawTestData)) == 0]
# bad columns
cleanTestData <- cleanTestData[, -c(1:7)]
str(cleanTestData)
```

## Data Slicing

Below is some code that splits the training data into training/validation datasets:

```{r}
library(caret);
set.seed(1234) 
inTrain <- createDataPartition(y = cleanTrainData$classe, p = 0.7, list = FALSE)
training <- cleanTrainData[inTrain,]
str(training)
validation <- cleanTrainData[-inTrain, ]
str(validation)
```

## Prediction Model Definition (Using Random Forest)

```{r}
library(doParallel)
model <- "fitRF.RData"
if (!file.exists(model)) {
    require(parallel)
    require(doParallel)
    cluster <- makeCluster(detectCores() - 1)
    registerDoParallel(cluster)
    control <- trainControl(method = "cv", number = 5)
    fitRF <- train(classe ~ ., data = training, method="rf", trControl = control)
    save(fitRF, file = "fitRF.RData")
    stopCluster(cluster)
} else { 
    load(file = "fitRF.RData")
}
```

## Accuracy and Sample Error of Model

### Training
```{r}
pred0 <- predict(fitRF, training)
confusionMatrix(pred0, training$classe)
```

### Validation
```{r}
pred1 <- predict(fitRF, validation)
confusionMatrix(pred1, validation$classe)
```
Based upon the results from the validation, the accuracy of the model is very high at .9947.

### Final Model
```{r}
fitRF$finalModel
```

## Use model on test data
```{r}
predTest <- predict(fitRF, cleanTestData)
predTest
```

## Conclusion

The model predicted the final 20 test cases quiz with 100% accuracy.

